Network Cascades

### 2016-09-22 paper/paper.tex
Incorporate review 2's suggestions

In this configuration, nothing actually occurs yet however, since all states are $0$. All agents do periodically check their neighbors and calculate if they should change their own state, but of course this always results in no change. To actually begin the simulation properly, we must create the initial impulse. That is, we change the state of one node to $1$ at a random point in time. In this case, we have chosen the node with threshold $\phi = 1$. The new state can be seen in figure \ref{fig:model2}.

@@

As the simulation keeps going, eventually the neighboring node with $\phi = 0.1$ will inspect its neighbors and notice the change. It has $4$ neighbors, of which $1$ is in state $1$. That means it compares $\frac{1}{4}$ to its own threshold of $0.1$. Since the calculated value is above the threshold, the node will change its own state to $1$. The new state of the network can be seen in figure \ref{fig:model3}.

@@

Next, we will look at the node with threshold $\phi = 0.5$. It has $3$ neighbors, of which $2$ have already changed their state. Since $\frac{2}{3} > 0.5$, the node also changes its state and the graph's state looks like in figure \ref{fig:model4}.

@@

As should be obvious by now, the final two nodes will now also change their state. First the node with $\phi = 0.5$ and finally the node with $\phi = 0.8$. At that point, the entire network is in a stable, unchanging state, as can be seen in figure \ref{fig:model5}. Since the entire network has changed its state to a uniform $1$, the cascade successfully worked its way from a single impulse to covering the entire network.

@@

The cascade window diagram is limited to a binary statement whether cascades are possible. To gain further insight on  the nature of cascades inside the window, Watts examined ``cross section'' of the window for $\phi=0.18$ (marked in figure \ref{fig:cascade-window} with by the densely stroked line at $\phi=0.18$), consisting of several measurements. They are depicted in figure \ref{fig:cross-section}. The upper graph (a) maps from $z$ to the amount of steps required for the cascade to occur. The lower figure (b) maps from $z$ to various percentages.

@@

Further insights on the cascade window can be gained by examining the size of cascades around its boundaries. Figure \ref{fig:cascade-size} shows the cumulative distribution of the cascade sizes. The size of a cascade is measured in terms of the ratio between agents in state $1$ to the total number of agents. The examined section of the window is the same as in figure \ref{fig:cross-section}. See \ref{sec:findings} for Watts' interpretation of this graph.

@@

In the above figure \ref{fig:cross-section} was introduced as a means to compare the properties of the cascade window for a given threshold. Watts' interpretation of the results is as follows:

@@

The average size of global cascades (solid line and solid circles, the upper graphs) is obviously not related to the size of the vulnerable cluster or the extended vulnerable cluster, since it does not decline with decreasing sizes for greater values of $z$. It is only governed by the $z$, i.e. the connectivity of the whole network. The relationship between connectivity and cascade size is further examined below, using figure \ref{fig:cascade-size}. The intuition that with decreasing size of the vulnerable cluster (to which the cascade is limited) the size of cascades should also decrease is false. Although unable to give a definitive explanation, Watts tries to explain this phenomenon: In the later stages of the cascade, the cascade has switched a lot of vulnerable nodes to $1$. In this situation it is possible that stable nodes, i.e. nodes that need more than one neighbor switching to switch themselves, have multiple switched adjacent vulnerable nodes and now do switch. This process could repeat itself until the entire connected component to which the vulnerable cluster belongs might be switched. Watts did not investigate in this paper if this phenomenon only occurs in the examined class of random graphs or if it also appears in other random graphs.

@@

The same goes for a change in connectivity distribution, as can be seen in figure \ref{fig:dunnolol}. A uniform Poisson distribution has a higher rate of cascades than a power-law distributed graph. This can be explained by the compound effects from before. In a power law distribution, there are pockets of low connectivity, where it impedes the progress of the cascade. There are also pockets of high connectivity, leading to nodes with many neighbors, which makes reaching their neighbor threshold difficult.

>>>>

In this configuration, nothing actually occurs yet however, since all states are $0$. All agents do periodically check their neighbors and calculate if they should change their own state, but of course this always results in no change. To actually begin the simulation properly, we must create the initial impulse. That is, we change the state of one node to $1$ at a random point in time. In this case, we have chosen the node with threshold $\phi = 1$. The new state can be seen in Figure \ref{fig:model2}.

@@

As the simulation keeps going, eventually the neighboring node with $\phi = 0.1$ will inspect its neighbors and notice the change. It has $4$ neighbors, of which $1$ is in state $1$. That means it compares $\frac{1}{4}$ to its own threshold of $0.1$. Since the calculated value is above the threshold, the node will change its own state to $1$. The new state of the network can be seen in Figure \ref{fig:model3}.

@@

Next, we will look at the node with threshold $\phi = 0.5$. It has $3$ neighbors, of which $2$ have already changed their state. Since $\frac{2}{3} > 0.5$, the node also changes its state and the graph's state looks like in Figure \ref{fig:model4}.

@@

As should be obvious by now, the final two nodes will now also change their state. First the node with $\phi = 0.5$ and finally the node with $\phi = 0.8$. At that point, the entire network is in a stable, unchanging state, as can be seen in Figure \ref{fig:model5}. Since the entire network has changed its state to a uniform $1$, the cascade successfully worked its way from a single impulse to covering the entire network.

@@

The cascade window diagram is limited to a binary statement whether cascades are possible. To gain further insight on  the nature of cascades inside the window, Watts examined ``cross section'' of the window for $\phi=0.18$ (marked in Figure \ref{fig:cascade-window} with by the densely stroked line at $\phi=0.18$), consisting of several measurements. They are depicted in Figure \ref{fig:cross-section}. The upper graph (a) maps from $z$ to the amount of steps required for the cascade to occur. The lower figure (b) maps from $z$ to various percentages.

@@

Further insights on the cascade window can be gained by examining the size of cascades around its boundaries. Figure \ref{fig:cascade-size} shows the cumulative distribution of the cascade sizes. The size of a cascade is measured in terms of the ratio between agents in state $1$ to the total number of agents. The examined section of the window is the same as in Figure \ref{fig:cross-section}. See \ref{sec:findings} for Watts' interpretation of this graph.

@@

In the above Figure \ref{fig:cross-section} was introduced as a means to compare the properties of the cascade window for a given threshold. Watts' interpretation of the results is as follows:

@@

The average size of global cascades (solid line and solid circles, the upper graphs) is obviously not related to the size of the vulnerable cluster or the extended vulnerable cluster, since it does not decline with decreasing sizes for greater values of $z$. It is only governed by the $z$, i.e. the connectivity of the whole network. The relationship between connectivity and cascade size is further examined below, using Figure \ref{fig:cascade-size}. The intuition that with decreasing size of the vulnerable cluster (to which the cascade is limited) the size of cascades should also decrease is false. Although unable to give a definitive explanation, Watts tries to explain this phenomenon: In the later stages of the cascade, the cascade has switched a lot of vulnerable nodes to $1$. In this situation it is possible that stable nodes, i.e. nodes that need more than one neighbor switching to switch themselves, have multiple switched adjacent vulnerable nodes and now do switch. This process could repeat itself until the entire connected component to which the vulnerable cluster belongs might be switched. Watts did not investigate in this paper if this phenomenon only occurs in the examined class of random graphs or if it also appears in other random graphs.

@@

The same goes for a change in connectivity distribution, as can be seen in Figure \ref{fig:dunnolol}. A uniform Poisson distribution has a higher rate of cascades than a power-law distributed graph. This can be explained by the compound effects from before. In a power law distribution, there are pockets of low connectivity, where it impedes the progress of the cascade. There are also pockets of high connectivity, leading to nodes with many neighbors, which makes reaching their neighbor threshold difficult.
